---
title: "Design a stroop effect test"
author: "SongTang"
date: "9/16/2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Backgroung Information
In a Stroop task, participants are presented with a list of words, with each word displayed in a color of ink. The participant’s task is to say out loud the color of the ink in which the word is printed. 
The task has two conditions: a congruent words condition, and an incongruent words condition. 

- In the congruent words condition, the words being displayed are color words whose names match the colors in which they are printed: for example RED, BLUE. 
- In the incongruent words condition, the words displayed are color words whose names do not match the colors in which they are printed: for example PURPLE, ORANGE. 

In each case, we measure the time it takes to name the ink colors in equally-sized lists. Each participant will go through and record a time from each condition.


**1. What is our independent variable? What is our dependent variable? **

**Answer:**

The independent variable is the congruency condition and incongruency condition - whether the name of the color matches with the ink color or not. The dependent variable is the time it takes to name the ink colors in equally-sized lists.

**2.  What is an appropriate set of hypotheses for this task? What kind of statistical test do you expect to perform? Justify your choices.**

**Answer:**


**3. Report some descriptive statistics regarding this dataset. Include at least one measure of central tendency and at least one measure of variability.**

```{r}
# Read in the data
library(readr)
dat <- read.csv("stroopdata.csv")
# Tidy up the data for later analysis
library(tidyr); suppressMessages(library(dplyr)) 
# Add a column identifying subjects
dat.subject <- mutate(dat, subject = 1:nrow(dat))
# Tidy up data by keeping one variable in one column
tidy.dat <- gather(dat.subject, congruency, time, -subject)
# Calculate the average time for both groups
tidy.dat %>%
    group_by(congruency) %>%
    summarise(round(mean(time), 2), round(median(time), 2), round(sd(time), 2), round(var(time), 2))
```


